Democratizing Secure Two-Party Computation

### 2017-10-12 OaklandSubmission/eval.tex
1.1
accuracy
eval 6.1

\section{Evaluation}
\label{sec:eval}
We evaluate \tool for the problem of {\it secure prediction}, where a server has a machine learning model, a client has an input, and after the protocol execution the client learns a label on the input.
From the protocol execution, the server learns nothing about the client input and the client learns nothing about the server model than what is revealed beyond the computed label.
This setting has been previously studied in~\cite{shafindss,minionn}
We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool to compare gainst 
previous work.
Next, we show the generality of \tool by implementing state-of-the-art machine learning models.
Finally, we discuss a model that requires pipelinig.
Our results are in two settings, a LAN setting and a cross-continent WAN setting.
The round trip time for two machines in a LAN setting is 1ms and in WAN setting is 40ms.
Each machine has a Intel(R) Xeon(R) CPU E5-2673 v3 processor running at 2.40GHz with
28 GBs of RAM. The \tool compiler is written in Python and compiles each of our benchmarks
in under a second to C++ code that makes calls to the ABY library~\cite{aby}. Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning next.

\subsection{Benchmarks}
In this paper, we focus on machine learning models for {\it classification}.
A {\it classifier} $C$ uses a trained {\it model} to {\it predict} a {\it label} $\ell$ for an input data point $x$. For example, given a data point which is a tuple of height and weight of an individual,
a classifier can predict a label ``male'' or ``female''.

A binary linear classifier is one of the simplest classifers. Here, the input data point $x\in\R^d$
and the model is a vector $w\in\R^d$. The possible labels $\ell\in\{\mathit{true},\mathit{false}\}$
and the classifier $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose operator.
This classifier requires $d$ multiplications, $d-1$ additions, and a comparison.

A more interesting classifier is Naive Bayes that predicts labels from the set $[n]$.
Here, the input data point is a {\it feature}
vector $x=(x_1,x_2,\ldots,x_d)^T$ where each $x_i\in [F]$.
The model has two matrices: a vector $P$ with $|P|=n$ and $P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood that the output label is $i$ for $i\in[n]$.
The other matrix $T$ has size $n\times d\times F$ and $T(i)(j)(x_j)\equiv\log p(x_j|\ell=i)$, i.e., the log-probability that 
the $j^{th}$ input feature is $x_j$ (where $j\in [d]$) given that the output label is $i$.
The classifier $C_{P,T}(x)$ outputs
\[
{\it argmax}_i P(i)+\sum_{j=1}^d T(i)(j)(x_j)
\]
This classifier requires $ndF$ comparisons and additions.

A decision tree of size $N$ takes as input $x\in\R^n$ and the model consists of a tree with a boolean predicate
assigned to each internal node. In particular, a node $j\in[N]$ at depth $i$ has a predicate $b_{j}^i\equiv x_i\leq w_i$. We start evaluating the tree from the root and if the predicate at the current node $b_j^i$ is false (true) then $x$ is passed to the left (right) node $b_{2j+1}^{i+1}$ ($b_{2j+2}^{i+1}$). This process is repeated till we reach a leaf. The leaves of the tree are labels and the output label is the leaf visited by this traversal.
Such a (binary) decision tree can be encoded as a polynomial and then prediction
reduces to polynomial evaluation. E.g., a binary tree with depth one and size $N=3$ can be encoded as the polynomial
$(1-b(x))\ell_1 + b(x)\ell_2$ where $b(x)$ is the predicate at the root, $\ell_1$ is the label of the left leaf and $\ell_2$ is the label of the right leaf.

The next class of classifiers that we discuss are deep neural nets or DNNs. 
A DNN has multiple layers that each compute a matrix multiplication followed by
an {\it activation} function. The most common activation functions are 
squaring $f(x)=x^2$ and ReLU $f(x)=\mathit{max}(x,0)$. 
If $A$ is a matrix then the output of $f(A)$ is also a matrix that contains $f$  applied
 to each entry of $A$ pointwise.
Given an input vector $x$, the predicted label of a DNN is
\[
 \mathit{argmax} W_n\cdot f_{N-1}(\ldots f_1(W_1\cdot x)\ldots)
\]
Here, $W_i$'s are matrices that are part of the model and $f_i$'s are the activation functions.
The operator $\cdot$ denotes a matrix multiplication.
Neural nets usually have one or more fully connected layers.
These layers multiply a matrix with a vector.
Some neural nets have convolution layers and such DNNs are also called CNNs.
For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication. The size of matrices manipulated by a convolution layer grows linearly with {\it window size} (typically 9 or 25), the number of {\it output channels} (typically 16, 32, or 64),
and the size of the matrix input to this layer. 
Therefore, fully connected layers are easier computation-wise compared
to convolution layers. In general, DNNs are computationaly heavy but provide
much better accuracies on computer vision tasks than the classifiers discussed above.

Finally, there are a class of machine learning classifiers that are much more efficient than
DNNs and provide reasonably good accuracies on standard learning tasks. One such classifer
is \bonsai~\cite{bonsai}, which is a variant of the decision trees described above.
In particular, each node $j$ evaluates a predicate $\theta_j^TZx > 0$ to decide whether
to pass $x$ to the left branch or the right branch ($\theta_j$ is a vector and $Z$ is a matrix).
The predicted value is
\[
{\it argmax} \sum_j I_j(x)W_j^TZx\circ f(V_j^TZx) 
\]
Here, $I_j$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
and is zero otherwise. 
The operation $\circ$ is a pointwise dot product, $W_k$'s and $V_k$'s are vectors,  and
the function $f$ is given by $f(x) = x$ if $|x| < 1$ and $\mathrm{signum}(x)$ otherwise.

@@

The results for linear classification are in Table~\ref{tab:lc}. The input and model are both vectors of length $d$. The time in columns ``Prev. time'' and communication ``Prev.Comm'' is the time and total network communication reported in~\cite{shafindss}. They remark that these times are obtained by simulating a network with 40ms round trip time, which is the same as our WAN setting. The total execution time of \tool generated code in the LAN and the WAN setting is reported next followed by the total communication.
The execution time is dominated by multiplications and boolean-ands.
The number of boolean-and gates (\#And) remains constant for both the benchmnarks as they reflect the number of and gates required for a single 32-bit integer comparison with zero. The number of multiplication gates (\#Mul) is the same as the model size $d$. The last column shows the lines of code of \tool program, which is also independent of the model size.

The results for Naive Bayes are in Table~\ref{tab:nb}. As before, $n$ denotes the length of the input feature vector $x=(x_1,\ldots,x_n)$ and $F$ is the number of possible values of each $x_i$.
Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons that result in many boolean-and gates (13k/750k denote 13000/75000). 
Here, again columns 4 and 5 show the results reported in~\cite{shafindss}. We observe that \tool generated code has competitive performance on this task. This improvement is significant as~\cite{shafindss} uses custom designed protocols and we are executing a generic two party computation (2PC). Moreover,~\cite{shafindss} remarks that generic 2PC did not scale to the smallest of their Naive Bayes classifiers and they had to scale down the prediction task and there a Yao-based implementation was 500 times slower. We have shown that by using \tool, generic 2PC can scale to real prediction tasks and provide performance competitive with specialized protocols.

This claim is further validated by Table~\ref{tab:dt}.
In~\cite{wu}, Wu et al. describe specialized protocols for trees and forests (that perfrom better than~\cite{shafindss}) and we report their performance in columns 3 and 4 of Table~\ref{tab:dt}.
Similar to~\cite{shafindss},~\cite{wu} remarks that these computations are infeasible with Yao-based 2PC.
Again, we observe that the performance of \tool generated generic 2PC code is competitive with specialized protocols. Therefore, we believe that \tool generated code can be a better baseline than Yao-based 2PC when comparing against generic 2PC-based implementations for secure prediction tasks.

@@

\subsection{Practical models}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of digits 0 to 9.
The first model that the tutorial describes is softmax regression
that provides an accuracy of 92\%. The classifier here evaluates
\[
 \mathit{argmax}_{i} w_i^Tx+b_i
\]
Here, $i\in[0,9]\cap\mathbb{Z}$, $x$ is a 784 length vector obtained from the input image,
each $w_i$ is a 784 length vector, and each $b_i$ is a scalar. 
We implement this classifer in \tool and the results are shown in the first row of Table~\ref{tab:tf}.
Since, we are not familiar with any other tools that have used this model as a benchmark,
we only report execution time of \tool generated code.
The column ``Model size'' is the number of 32 bit integers in the trained model.

>>>>

\section{Evaluation}
\label{sec:eval}
We evaluate \tool for the problem of {\it secure prediction}, where a server has a machine learning model, a client has an input, and after the protocol execution the client learns a label on his input.
From the protocol execution, the server learns nothing about the client's input and the client learns nothing about the server's model beyond what is revealed from the computed label.
This setting has been previously studied in~\cite{shafindss,minionn}
We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool to compare against 
previous work.
Next, we demonstrate the generality of \tool by implementing state-of-the-art machine learning models.
Finally, we discuss a model that requires pipelinig.
Our results are in two settings, a LAN setting and a cross-continent WAN setting.
The round trip time for two machines in a LAN setting is 1ms and in WAN setting is 40ms.
Each machine has a Intel(R) Xeon(R) CPU E5-2673 v3 processor running at 2.40GHz with
28 GBs of RAM. The \tool compiler is written in Python and compiles each of our benchmarks
in under a second to C++ code that makes calls to the ABY library~\cite{aby}. Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning and describe our benchmarks next.

\subsection{Benchmarks}
\divya{notation can be moved somewhere... } \\

\noindent\textbf{Notation.} We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. \rs{Define $\mathsf{argmax}$.}

In this paper, we focus on machine learning models for {\it classification}.
A {\it classifier} $C$ uses a trained {\it model} to {\it predict} a {\it label} $\ell$ for an input data point $x$. For example, given a data point which is a tuple of height and weight of an individual,
a classifier can predict a label ``male'' or ``female''. The {\it accuracy} of a classifier
refers to the fraction of data points that the classifier labels correctly from a given set of {\it test} data points.

A binary linear classifier is one of the simplest classifiers. Here, the input data point $x\in\R^d$
and the model is a vector $w\in\R^d$. The possible labels are $\ell\in\{\mathit{true},\mathit{false}\}$
and the classifier is $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose operator.
This classifier requires $d$ multiplications, $d-1$ additions, and a comparison.

A more interesting classifier is Na\"{i}ve Bayes that predicts labels from the set $[n]$.
Here, the input data point is a {\it feature}
vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
The model has two matrices: a vector $P$ of length $n$ and for each $i \in [n]$, $P(i)$ is the log-probability that the output label is $i$.
%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood that the output label is $i$ for $i\in[n]$.
The other matrix $T$ has size $n\times d\times F$ and for each $i\in [n], j\in [d], k\in [F]$, $T(i)(j)(k)$ is
%\equiv\log p(x_j|\ell=i)$, i.e., 
the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
The classifier $C_{P,T}(x)$ outputs
\[
{\sf argmax}_i P(i)+\sum_{j=0}^{d-1} T(i)(j)(x_j)
\]
This classifier requires $ndF$ comparisons and additions. \divya{Note that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix $T$  based on secret input $x_j$ resulting in $F$ comparisons (see \sectionref{}).} \cmmt{Is there a +1 etc missing or is this exact?}

A decision tree of size $N$ takes as input $x\in\R^d$ and the model consists of a binary tree of depth $d$ with a boolean predicate
assigned to each internal node. The root note is the $0^{th}$ node and for an internal node $j \in [N]$, the children nodes are $2j+1$ and $2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate $b_{j}^i\equiv x_i\leq w_i$. We start evaluating the tree from the root and if the predicate at the current node $b_j^i$ is false (resp., true) then $x$ is passed to the left (resp., right) child with predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is repeated till we reach a leaf. The leaves of the tree are labels and the output label is the leaf visited by this traversal.
Such a (binary) decision tree can be encoded as a polynomial (of degree linear in depth of the tree) and then the prediction task
reduces to polynomial evaluation. E.g., a binary tree with depth one and size $N=3$ can be encoded as the polynomial
$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label of the right leaf.

The next class of classifiers that we discuss are deep neural nets or DNNs. 
A DNN has multiple layers such that each layer computes a matrix multiplication followed by an {\it activation} function $f$. The most common activation functions are square $f(x)=x^2$ and rectifier linear unit (ReLU) $f(x)=\mathsf{max}(x,0)$. 
If $A$ is a matrix then the output of $f(A)$ is also a matrix that is obtained by applying $f$ to each entry of $A$ pointwise.
Given an input vector $x$, the predicted label of a DNN is
\[
 \mathsf{argmax}\ W_N\cdot f_{N-1}(\ldots f_1(W_1\cdot x)\ldots)
\]
Here, $f_i$'s are the \divya{public} activation functions and the model consists of matrices $W_i$. The input vector is $x \in \R^d$.
The operator $\cdot$ denotes a matrix multiplication.
Neural nets usually have one or more fully connected layers.
These layers multiply a matrix with a vector.
Some neural nets have convolution layers and such DNNs are also called convolutional neural nets or CNNs.
For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication. The size of matrices manipulated by a convolution layer grows linearly with {\it window size} (typically 9 or 25), the number of {\it output channels} (typically 16, 32, or 64), and the size of the matrix input to this layer. 
Therefore, fully connected layers are lighter computation-wise compared
to convolution layers. In general, DNNs are computationaly heavy but provide
much better accuracies on computer vision tasks than the classifiers discussed above.

Finally, there are a class of machine learning classifiers that are much more efficient than
DNNs and provide reasonably good accuracies on standard learning tasks. One such classifier
is \bonsai~\cite{bonsai} that takes as input $x \in \R^d$ and model consists of a binary tree with $N$ nodes. Each node $j$ contains matrices $W_j$ and $V_j$. The internal node $j$  evaluates a predicate $\theta_j^TZx > 0$ to decide whether
to pass $x$ to the left child $2j+1$ or the right child $2j+2$ (where $\theta_j$ is a vector and $Z$ is a matrix).
The predicted value is
\[
{\sf argmax} \sum_{j\in [N]} I_j(x)W_j^TZx\circ f(V_j^TZx) 
\]
Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
and is zero otherwise. 
The operation $\circ$ is a point-wise dot product, $W_j$'s and $V_j$'s are matrices. The activation function $f$ is given by $f(y) = y$ if $-1 < y < 1$ and $\mathrm{sign}(x)$ otherwise. The model consists of the matrices $W_j$, $V_j$, $Z$ and vectors $\theta_j$. \bonsai can be seen as a variant of decision trees where the prediction is a function of the path traversed from root to the leaf and not just the leaf itself.

@@

The results for linear classification are in Table~\ref{tab:lc}. The input and model are both vectors of length $d$. The time in columns ``Prev. time'' and communication ``Prev.Comm'' is the time and total network communication reported in~\cite{shafindss}. They remark that these times are obtained by simulating a network with 40ms round trip time, which is the same as our WAN setting. The total execution time of \tool generated code in the LAN and the WAN setting is reported next followed by the total communication.
The execution time is dominated by multiplications and boolean-ands.
The number of boolean-and gates (\#And) remains constant for both the benchmarks as they reflect the number of and gates required for a single 32-bit integer comparison with zero. The number of multiplication gates (\#Mul) is the same as the model size $d$. The last column shows the lines of code of \tool program, which is also independent of the model size.

The results for Na\"{i}ve Bayes are in Table~\ref{tab:nb}. As before, $n$ denotes the length of the input feature vector $x=(x_1,\ldots,x_n)$ and $F$ is the number of possible values of each $x_i$.
Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons that result in many boolean-and gates (13k/750k denote 13000/75000). 
Here, again columns 4 and 5 show the results reported in~\cite{shafindss}. We observe that \tool generated code has competitive performance on this task. This improvement is significant as~\cite{shafindss} uses custom designed protocols and we are executing a generic two party computation (2PC). Moreover,~\cite{shafindss} remarks that generic 2PC did not scale to the smallest of their Naive Bayes classifiers and they had to scale down the prediction task and there a Yao-based implementation was 500 times slower. We have shown that by using \tool, generic 2PC can scale to real prediction tasks and provide performance competitive with specialized protocols.

This claim is further validated by Table~\ref{tab:dt}.
In~\cite{wu}, Wu et al. describe specialized protocols for trees and forests (that perform better than~\cite{shafindss}) and we report their performance in columns 3 and 4 of Table~\ref{tab:dt}.
Similar to~\cite{shafindss},~\cite{wu} remarks that these computations are infeasible with Yao-based 2PC.
Again, we observe that the performance of \tool generated generic 2PC code is competitive with specialized protocols. Therefore, we believe that \tool generated code can be a better baseline than Yao-based 2PC when comparing against generic 2PC-based implementations for secure prediction tasks.

@@

\subsection{Practical models}
Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
Its introductory tutorial describes two prediction models for handwritten digit recognition
using the MNIST dataset~\cite{mnist}.
Each image in this dataset is a greyscale $28\times 28$ image of digits 0 to 9.
The first model that the tutorial describes is softmax regression
that provides an accuracy of 92\%. The classifier here evaluates
\[
 \mathsf{argmax}_{i} w_i^Tx+b_i
\]
Here, $i\in[0,9]\cap\mathbb{Z}$, $x$ is a 784 length vector obtained from the input image,
each $w_i$ is a 784 length vector, and each $b_i$ is a scalar. 
We implement this classifer in \tool and the results are shown in the first row of Table~\ref{tab:tf}.
Since, we are not familiar with any other tools that have used this model as a benchmark,
we only report execution time of \tool generated code.
The column ``Model size'' is the number of 32 bit integers in the trained model.
